---
title: "Lab 3: Reproducible report"
author: "Nikita Makarevich 153989, Szymon Siemieniuk 151***"
date: "2023-04-15"
output: 
  html_document:
    toc: true 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyr)
library(ggplot2)
library(plotly)

invents = read.csv("rebrickable/inventories.csv.gz")

invent_sets = read.csv("rebrickable/inventory_sets.csv.gz")
sets = read.csv("rebrickable/sets.csv.gz")
themes = read.csv("rebrickable/themes.csv.gz")

invent_minifigs = read.csv("rebrickable/inventory_minifigs.csv.gz")
minifigs = read.csv("rebrickable/minifigs.csv.gz")

invent_parts = read.csv("rebrickable/inventory_parts.csv.gz")
parts = read.csv("rebrickable/parts.csv.gz")
colors = read.csv("rebrickable/colors.csv.gz")
elements = read.csv("rebrickable/elements.csv.gz")
part_categs = read.csv("rebrickable/part_categories.csv.gz")
part_rels = read.csv("rebrickable/part_relationships.csv.gz")

```

# **Rebrickable**. Data Exploration

_Rebrickable_ has been so kind as to provide us with data on their inventories and all related attributes. So we decided to take some time to practice our visualization skills on real data (as well as maybe find interesting patterns in the data)


## History of Rebrickable with your own eyes

_Rebrickable_ is a company, therefore we would expect it to grow with time (assuming it is successful). But does the data confirm this? Even though we don't have numbers on the revenue of this company :(, we can at least see how many sets were put out with time: 

```{r historical, echo=FALSE, warning=FALSE, message=FALSE}
years_set_amnts = sets %>%
  group_by(year) %>%
  summarize(n=n(), med_num_parts=median(num_parts))

years_set_amnts$num_parts_cat = 
  cut(years_set_amnts$med_num_parts, breaks=c(0, 20, 40, 60, 80, 100, +Inf))

ggplot(years_set_amnts, aes(x=year, y=n, fill=factor(num_parts_cat))) +
  geom_bar(stat="identity", color="#000000") +
  scale_fill_brewer(palette="YlGnBu") + 
  stat_smooth(formula=y~x, fill=NA, alpha=0.3, size=1.3) +
  scale_y_continuous(limits=c(0, 1250), expand=c(0,0)) +
  labs(fill="Median number \nof parts in set",
       title="Number of new sets by year",
       y="Number of sets") + 
  theme_minimal() +
  theme(plot.title=element_text(hjust=0.5))
```

## *Rebrickable* doesn't see color

```{r color_counts, echo=FALSE}
categ_color_cnts = invent_parts %>%
  inner_join(parts, by="part_num") %>%
  group_by(part_cat_id) %>%
  summarize(color_cnt=n_distinct(color_id)) %>%
  inner_join(part_categs, by=join_by(part_cat_id == id)) %>%
  arrange(color_cnt)

splitcat = rep("middle", nrow(categ_color_cnts))
n_best = 10
splitcat[1:n_best] = rep(paste("Worst", n_best), n_best)
splitcat[(66-n_best+1):66] = rep(paste("Best", n_best), n_best)
categ_color_cnts$splitcat = splitcat
categ_color_cnts = categ_color_cnts %>%
  filter(splitcat != "middle")

bar_colors = rep(c("#225ea8", "#41b6c4"), length.out=nrow(categ_color_cnts))

p = ggplot(categ_color_cnts, aes(x=factor(name, levels=name), y=color_cnt)) +
  geom_bar(stat="identity", fill=bar_colors) +
  coord_flip() +
  scale_y_continuous(limits=c(0, 120), expand=c(0, 0)) +
  facet_grid(splitcat ~ ., scales="free") +
  labs(y="Number of available colors",
       x="Category",
       title="Available colors for each category") +
  theme_minimal()

p
```

## Top 10 Hot Themes in Your Area!! <3

Millions of people around the world build sets specifically because of the theme (perhaps their favorite show or movie). So we decided to see which themes have been most abundant in the inventories of *Rebrickable* and when did they first appear!

```{r hot_themes, echo=FALSE}
themes_first_years = sets %>%
  group_by(theme_id) %>%
  summarize(first_year=min(year))

themes_quants = invent_sets %>%
  inner_join(sets, by="set_num") %>%
  group_by(theme_id) %>%
  summarize(tot_quant=sum(quantity)) %>%
  top_n(10, tot_quant)

theme_avg_parts = sets %>%
  group_by(theme_id) %>%
  summarize(avg_parts=mean(num_parts))

thm_quants_fyears = themes_first_years %>%
  inner_join(themes_quants, by="theme_id") %>%
  inner_join(theme_avg_parts, by="theme_id") %>%
  inner_join(themes, by=join_by(theme_id == id)) %>%
  arrange(desc(avg_parts))

colors = c("#081d58", "#253494", "#225ea8", "#1d91c0", "#41b6c4",
           "#7fcdbb", "#c7e9b4", "#edf8b1", "#ffffd9", "#ffffff")
vjusts = c(3.6, 4, 3.3, 3, 3, 2, -2.5, 1.7, 1, -2)
hjusts= c(-0.2, 0.5, 0.5, 0.5, 0.5, 0.5, 0.5, 0.5, 0.5, 0.5)
ggplot(thm_quants_fyears, aes(x=first_year, y=tot_quant,label=name)) +
  geom_point(aes(size=avg_parts), fill=colors, color="black", shape=21, alpha=0.7) +
  scale_size(name="Size", range=c(5, 35)) +
  geom_point(size=1.5, color="black") +
  geom_text(size=4, color="black", hjust=hjusts, vjust=-vjusts) +
  scale_x_continuous(limits=c(1980, 2025)) +
  scale_y_continuous(expand=c(0.1, 0.1)) +
  theme_minimal() +
  theme(legend.position="none", plot.title=element_text(size=18, hjust=0.5)) +
  labs(x="Year of First Introduction",
       y="Total Quantity in Inventories",
       title="Top 10 Themes of Sets",
       subtitle="(size indicates average number of parts in set)")
```